---
title: "Readind XML into R"
author: "Yimi Zhao"
date: "Tuesday, May 05, 2015"
output: html_document
---

## Simple Tutorial for Readind XML into R

If you want to know more about XML and XPath, check the following tutorials:

   * XML Tutorial <http://www.w3schools.com/xml>
   * XPath Tutorial <http://www.w3schools.com/xpath>

### In R, we apply "XML" package to read XML files
```{r, message=FALSE}
#install the package
# install.packages("XML")
library(XML)

fileUrl<-"http://www.w3schools.com/xml/cd_catalog.xml"
(doc<-xmlTreeParse(fileUrl,useInternal=T))
```

### Check the imported XML file:

```{r}
# check the xml file
rootNode<-xmlRoot(doc)
xmlName(rootNode) # root element name
length(names(rootNode))
```
```{r}
rootNode[[1]]
rootNode[[1]][[1]]

```

### Using xmlSAaaply() to extract element values through XPath syntax:

```{r}
# extract values through XPath
xmlSApply(rootNode,xmlValue)

TITLE<-xpathSApply(rootNode,"//TITLE",xmlValue)
ARTIST<-xpathSApply(rootNode,"//ARTIST",xmlValue)
COUNTRY<-xpathSApply(rootNode,"//COUNTRY",xmlValue)
#COMPANY<-xpathSApply(rootNode,"//COMPANY",xmlValue)
PRICE<-xpathSApply(rootNode,"//PRICE",xmlValue)
YEAR<-xpathSApply(rootNode,"//YEAR",xmlValue)

```

### Create a data frame based on extracted values we got from the XML file:

```{r}
# merge into a new data frame
CD_catalog<-data.frame(TITLE, ARTIST, COUNTRY, PRICE, YEAR)
dim(CD_catalog)
CD_catalog

```

### Finally, we can  write the tidy data-framed data into a csv file:

```{r}
#export the tidy data into a csv file
 #write(CD_catalog,"CD_catalog.csv")
```

